Follow-up Samples

### 2016-12-07 submission/manuscript_outline_20161024.tex

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a \texttt{66} person group that included samples before
and after treatment to allow for the assessment of how the model adjusts
risk after treatment. The model chosen had a 10-fold cross validated AUC
of \texttt{0.819} and a test set AUC of \texttt{0.815}. : For the follow
up samples our Random Forest model significantly decreased the positive
probability of lesion compared to the initial samples for both adenoma
(P-value = \texttt{7.95e-08}) and carcinoma (P-value =
\texttt{7.95e-08}). In the test set, the prediction of lesion for the
initial samples had a sensitivity of \texttt{100}\% (\texttt{66/66})
while follow up it was \texttt{100}\% (\texttt{1\ /\ 1}). Our model
predicted that \texttt{36.4}\% of the 67-person cohort had normal colons
and \texttt{63.6}\% had a lesion. Our model suggests that treatment does
significantly reduce the probability of having a colonic lesion. Further
surveillance of these individuals will enable us to determine whether
models such as the one we present here can also be used to predict
recurrence of colorectal cancer.

@@

\textbf{\emph{Bacterial Community and Fit Changes before and after
Treatment}} Based on thetayc distance metrics comparing the initial to
the follow up samples there was no difference between the adenoma and
carcinoma group (P-value = \texttt{0.697}) {[}Figure 1a{]}. There was a
difference in FIT between initial and follow up samples with the
carcinoma group having a significant decrease in FIT versus the adenoma
group (P-value = \texttt{2.15e-05}) {[}Figure 1b{]}. Although the
thetayc distance metric change was similar between adenoma and carcinoma
the directionality of the change was significant in the cancer group
between initial and follow up (P-value = \texttt{0.001}) but not for the
adenoma group (P-value = \texttt{1}) {[}Figure 2{]}. When all follow up
samples were compared to each other there was no significant overall
difference between them (P-value = \texttt{0.643}). There was no
significant difference between initial and follow up samples for
observed OTUs, Shannon diversity, and evenness after correction for
multiple comparisons {[}Table S1{]}. Time of follow up sample from
initial sampling did not have an effect on our prediction data set
(uncorrected P-value = \texttt{0.784}).

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of \texttt{0.795} to a maximum of
\texttt{0.854}. To be conservative the model chosen for prediction of
disease in the follow up samples had an AUC in the middle of all the
\texttt{100} runs which was \texttt{0.819}. Interestingly, the worst AUC
model from training performed the best on it's respective 80/20 split
test data {[}Figure 3{]}. In fact the 80/20 test performance showed that
the AUC for the middle model chosen was the most stable (best training
model test set AUC = \texttt{0.682}, middle training model test set AUC
= \texttt{0.815}, worse training model test set AUC = \texttt{0.931}).
That is to say it had the smallest change in AUC in comparison to the
other minimum and maximum AUC trained models. There was no significant
difference between the AUC of the best and middle training models
(P-value = \texttt{0.419}). There was also no difference in the middle
model versus worse (P-value = \texttt{0.178}) or full data model
(P-value = \texttt{1}). The two comparisons with a significant
difference were between the worse training model and the best training
model (P-value = \texttt{7.94e-04}) and the worse training model and
full data model (P-value = \texttt{3.52e-03}).

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of \texttt{32} variables identified as being present in
more than 50\% of the training models {[}Table S2{]}. The top 5 most
important bacterial OTUs were \texttt{Bacteria\ (Otu000013)},
\texttt{Escherichia/Shigella\ (Otu000018)},
\texttt{Bacteria\ (Otu000020)}, \texttt{Ruminococcus\ (Otu000017)}, and
\texttt{Porphyromonas\ (Otu000153)}. These 5 OTUs were present in at
least \texttt{99} out of the total \texttt{100} different 80/20 runs.

\textbf{\emph{Surgical Removal of an Adenoma or Carcinoma Results in a
Decrease in Positive Probability Prediction}} A total of \texttt{1}
sample was omitted from the original \texttt{67} sample set since it was
missing a complete set of follow up data. This left a total of
\texttt{66} samples for test predictions. After multiple comparison
correction there was a significant overall decrease in positive
probability of a carcinoma and adenoma (P-value = \texttt{1.14e-11})
{[}Figure 4{]}. This decrease was significant for both adenoma (P-value
= \texttt{7.95e-08} {[}Figure 4a{]} and carcinoma (P-value =
\texttt{7.95e-08}) {[}Figure 4b{]} alone. This also held specifically
for those with screen relative neoplasias (SRN) (P-value =
\texttt{5.35e-04}). A total of \texttt{66} or \texttt{100}\% of all
samples were correctly predicted to have an adenoma or carcinoma.
Although there was a decrease in positive probability only \texttt{24}
of the total \texttt{66} individuals were classified as adenoma or
carcinoma free on follow up (successful classification of
\texttt{37.9}\%). There was no significant difference between the
predictions and actual diagnosis for the initial samples in the
67-sample cohort test set. However, the predictions were significantly
discordent with the diagnosis for the follow up samples (P-value =
\texttt{4.19e-10}). Although there were discordent results the
respective sensitivity for the initial group was \texttt{100}\% and for
follow up was \texttt{100}\%, respectively.

There was \texttt{1} individual who still clearly had CRC on follow up
as well as \texttt{5} individuals whose status on follow up was unknown.
Although the \texttt{1} individual had a decrease in positive
probability their follow up sample was still higher than the cutoff
threshold of 0.5 (positive probability = \texttt{0.927}). Interestingly,
\texttt{2} individuals who were unknown on follow up still were over the
threshold cutoff of 0.5 even though, like the \texttt{1} individual with
clear CRC on follow up, the probability of an adenoma or carcinoma
decreased {[}Table S3{]}.

The follow up positive probabilities were not affected by either
chemotherapy treatment (uncorrected P-value = \texttt{0.919}) or
radiation therapy (uncorrected P-value = \texttt{1}). There was also no
difference in the amount of change in the positive probability based on
whether individuals received chemotherapy (unccorected P-value =
\texttt{0.578}) or radiation therapy (uncorrected P-value =
\texttt{0.904}).

\textbf{\emph{Specific OTUs in the Lesion Model are not Detected in
Follow Up Versus Initial Samples}} Overall, there were a total of
\texttt{8} OTUs that were common between the main lesion model and the
model for classifying initial and follow up samples specifically
{[}Table S4{]}. A total of \texttt{1} OTU was still significant after
multiple comparison correction and it's lowest taxonomic identification
was to \texttt{Blautia}. In general, \texttt{Otu000012}
(\texttt{Blautia}) was decreased from initial to follow up {[}Figure
5{]}. The relative abundance was not drastically different then the mean
of the values observed in the control training set {[}Figure 5{]}.

@@

We next created a model that incorporated both patient metadata, FIT,
and the bacterial microbiome to be able to predict lesions (adenoma or
carcinoma). We chose the middle training model, based on AUC, from 100
80/20 (train/test) splits. It's 10-fold cross validated AUC was similar
to it's test set AUC which was not the case for both the best and worse
training model {[}Figure 3{]}. Using this model we predicted the
probability of a lesion in the inital and follow up samples {[}Figure
4{]}. There was a signficant decrease in positive probability regardless
of whether the sample as a carcinoma or adenoma. The overall sensitivity
for lesion detection in the intial samples was \texttt{100} and for
follow ups was \texttt{100}. Although there was a decrease in overall
probability of an adenoma or carcinoma only \texttt{24} were below the
0.5 threshold out of the total \texttt{65} individuals who were
diagnosed as not having a carcinoma on follow up.

We then investigated which OTUs could potentially be more important in
our model {[}Figure 5 \& Table S4{]}. Only a single OTU was significant
after multiple comparison correction and the lowest taxonomic
identification of \texttt{Otu000012} was to \texttt{Blautia}. Although
there was a difference in the relative abundance at initial and follow
up these values were not drastically different from the relative
abundance values observed in the control individuals of the training set
{[}Figure 5{]}. This research provides evidence that it is possible to
create a highily sensitive model for detection of adenoma or carcinoma
with bacterial microbiome data. It accomplishes this by using a unique
sample set in which before and after stool samples are available for
assessment. By using these types of samples we are able to show that
this model is reactive. That is to say that after surgery for removal of
the adenoma or carcinoma it decreases the positive probability to
reflect a lower likelihood of the individual having an adenoma or
carcinoma.

@@

\textbf{\emph{Study Design and Patient Sampling}} The sampling and
design of the study was similar to that reported in Baxter, et al
(\textbf{insert citation}). In brief, study exclusion involved those who
had already undergone surgery, radiation, or chemotherapy, had
colorectal cancer before a baseline stool sample could be obtained, had
IBD, a known hereditary non-polyposis colorectal cacner, or Familial
adenomatous polyposis. Samples used to build the model used for
prediction were collected either prior to a colonoscopy or between 1 - 2
weeks after. The bacterial microbiome has been shown to nomralize within
this time period (\textbf{insert citation}). Kept apart from this
training set were a total of \texttt{67} individuals that not only had a
sample as described previoulsy but also a follow up sample between
\texttt{188} - \texttt{546} days after surgery and treatment had been
completed. This study was approved by the University of Michigan
Institutional Review Board. All study participants provided informed
consent and the study itself conformed to the guidelines set out by the
Helsinki Declaration.

@@

\textbf{\emph{Sequence Processing}} The mothur software package
(v1.37.5) was used to process the 16S rRNA gene sequences. This process
has been previously described (\textbf{insert citations}). The general
processing workflow using mothur is as follows: Paired-end reads were
first merged into contigs, quality filtered, aligned to the SILVA
database, screening for chimeras, classified with a naive Bayesian
classifier using the Ribosomal Database Project (RDP), and clustered
into Operational Taxonomic Units (OTUs) using a 97\% similarity cutoff
with an average neighbor clustering algorithm. The numer of sequences
for each sample was rarified to \texttt{10521} in an attempt to minimize
uneven sampling.

\textbf{\emph{Model Creation}} The Random Forest (\textbf{insert
citation}) algorithm was used to create the model used for prediction of
lesion (adenoma or carcinoma) for the \texttt{67} individuals with
follow up samples. The model included data on FIT, the bacterial
microbiome, sex, age, Body Mass Index (BMI), whether the indivdiual was
caucasion or not, history of cancer, and family history of cancer.
Non-binary data was checked for near zero variance and auto correlation.
Data columns that had near zero variance were removed. Columns that were
correlated with each other over a Spearman correlation coefficient of
0.75 had one of the two columns removed. This pre-processing was
performed with the R package caret (v\texttt{6.0.73}). Optimiation of
the mtry hyperparameter as well as data on the best and worse
performance of the model involved taking the \texttt{490} samples and
making 100 80/20 (train/test) splits in the data where control and
lesion were equally represented in the 80 and 20 split, respectively.
This 80\% portion was then split again into an 80/20 split, and run
twenty times through 10-fold cross validation to optimzie the model's
Receiver Operator Characteristic (ROC) and tuneing both mtry and ntree.
This resulting model was then tested on the 20\% of the data that was
originally held out from this overall process. This was repeated 100
times with the best model after all the repeats chosen to be used on the
\texttt{67} samples with follow up. Once the ideal mtry was found the
entire \texttt{490} sample set was used to create the final Random
Forest model. The default cutoff of a probability of 0.5 was used as the
threshold to classify individuals as positive or negative of lesion.

@@

\textbf{\emph{Statistical Analysis}} The R software package
(v\texttt{3.3.0}) was used for all statisitical analysis. Comparisons
between bacterial community structure utilized PERMANOVA (\textbf{insert
citation}) in the vegan package (v\texttt{2.4.1}) while comparisons
between ROC curves utilized the method by DeLong et al. (\textbf{insert
citation}) executed by the pROC (v\texttt{1.8}) package. Comparisons
between probabilities as well as overall amount of OTU between initial
and follow up samples utilized a paired wilcoxson ranked sum test. Where
multiple comparison testing was needed a Benjamini-Hochberg (BH)
correction was applied (\textbf{insert citation}) and a corrected
P-value of less than 0.05 was considered significant. Unless otherwise
stated the P-values reported are those of the BH corrected ones.

@@

\textbf{Figure 1: Change in Thetayc and Fit between initial and follow
up in adenoma or carcinoma group.} A) No significant difference was
found between the adenoma and carcinoma group for thetayc (P-value =
\texttt{0.697}). B) A significant difference was found between the
adenoma and carcinoma group for Fit (P-value = \texttt{2.15e-05}).

@@

\textbf{Figure 3: Graph of the Receiver Operating Characteristic Curve
on Test Set Performance of the Best, Middle, and Worse Training Models.}
For each of the 100 training cohort sets used had \texttt{392}
individuals and the testing cohort sets had \texttt{13} individuals. The
AUC on the test sets for the best, middle, and worse models from
training were \texttt{0.682}, \texttt{0.815}, and \texttt{0.931},
respectively. cvAUC is the 10-fold cross-validated AUC from training.

@@

\textbf{Figure 5: Lesion Model OTU with a Significant Decrease in
Relative Abundance that is also Predictive of Initial and Follow Up.}
After multiple comparison correction \texttt{1} (\texttt{Blautia}) was
the only one with a P-value \textless{} 0.05. The dotted line represents
the average relative abundance in the control training group.

>>>>

Colorectal cancer (CRC) continues to be a worldwide health problem with
early detection being used as a key component in mitigating deaths due
to the disease. Previous research suggests that the bacterial microbiome
can be used as a biomarker for colorectal cancer (CRC). These reports
have mostly focused on investigating how the bacterial microbiome is
used at a single point in time to predict disease. In this study, we
assessed whether a model built with bacterial microbiome data could
accurately predict adenoma or carcinoma and adjust positive probability
for lesion (adenoma or carcinoma) after surgical treatment. This model
was tested on a 66 person group that included samples before and after
treatment to allow for the assessment of how the model adjusts risk
after treatment. The model used for prediction had an AUC of 0.813. For
the follow up samples our Random Forest model significantly decreased
the positive probability of lesion compared to the initial samples for
both adenoma (P-value = 7.95e-08) and carcinoma (P-value = 7.95e-08).
Our model predicted that 36.4\% of the 67-person cohort had normal
colons and 63.6\% had a lesion. Some OTUs that changed the most before
and after treatment included OTUs that were affiliated with members of
Blautia, Clostridium\_XlVa and Escherichia/Shigella. Our model suggests
that treatment does significantly reduce the probability of having a
colonic lesion. Further surveillance of these individuals will enable us
to determine whether models such as the one we present here can also be
used to predict recurrence of colorectal cancer.

@@

\textbf{\emph{Bacterial Community and Fit Changes before and after
Treatment}} Based on thetayc distance metrics comparing the initial to
the follow up samples there was no difference between the adenoma and
carcinoma group (P-value = 0.697) {[}Figure 1a{]}. There was a
difference in FIT between initial and follow up samples with the
carcinoma group having a significant decrease in FIT versus the adenoma
group (P-value = 2.15e-05) {[}Figure 1b{]}. Although the thetayc
distance metric change was similar between adenoma and carcinoma the
directionality of the change was significant in the cancer group between
initial and follow up (P-value = 0.002) but not for the adenoma group
(P-value = 0.997) {[}Figure 2{]}. When all follow up samples were
compared to each other there was no significant overall difference
between them (P-value = 0.085). There was no significant difference
between initial and follow up samples for observed OTUs, Shannon
diversity, and evenness after correction for multiple comparisons
{[}Table S1{]}. Time of follow up sample from initial sampling did not
have an effect on our prediction data set (uncorrected P-value = 0.784).

\textbf{\emph{Outcome of Model Training}} The range of the AUC for model
training ranged from a minimum of 0.795 to a maximum of 0.854. To be
conservative the model chosen for prediction of disease in the follow up
samples had an AUC in the middle of all the 100 runs which was 0.819.
Interestingly, the worst AUC model from training performed the best on
it's respective 80/20 split test data {[}Figure 3{]}. In fact the 80/20
test performance showed that the AUC for the middle model chosen was the
most stable (best training model test set AUC = 0.682, middle training
model test set AUC = 0.815, worse training model test set AUC = 0.931).
That is to say it had the smallest change in AUC in comparison to the
other minimum and maximum AUC trained models. There was no significant
difference between the AUC of the best and middle training models
(P-value = 0.419). There was also no difference in the middle model
versus worse (P-value = 0.178) or full data model (P-value = 1). The two
comparisons with a significant difference were between the worse
training model and the best training model (P-value = 7.94e-04) and the
worse training model and full data model (P-value = 3.52e-03).

\textbf{\emph{Most Important Variables to the Model}} Overall, there
were a total of 32 variables identified as being present in more than
50\% of the training models {[}Table S2{]}. The top 5 most important
bacterial OTUs were Bacteria (Otu000013), Escherichia/Shigella
(Otu000018), Bacteria (Otu000020), Ruminococcus (Otu000017), and
Porphyromonas (Otu000153). These 5 OTUs were present in at least 99 out
of the total 100 different 80/20 runs.

\textbf{\emph{Surgical Removal of an Adenoma or Carcinoma Results in a
Decrease in Positive Probability Prediction}} A total of 1 sample was
omitted from the original 67 sample set since it was missing a complete
set of follow up data. This left a total of 66 samples for test
predictions. After multiple comparison correction there was a
significant overall decrease in positive probability of a carcinoma and
adenoma (P-value = 1.14e-11) {[}Figure 4{]}. This decrease was
significant for both adenoma (P-value = 7.95e-08 {[}Figure 4a{]} and
carcinoma (P-value = 7.95e-08) {[}Figure 4b{]} alone. This also held
specifically for those with screen relative neoplasias (SRN) (P-value =
5.35e-04). A total of 66 or 100\% of all samples were correctly
predicted to have an adenoma or carcinoma. Although there was a decrease
in positive probability only 24 of the total 66 individuals were
classified as adenoma or carcinoma free on follow up (successful
classification of 37.9\%). There was no significant difference between
the predictions and actual diagnosis for the initial samples in the
67-sample cohort test set. However, the predictions were significantly
discordent with the diagnosis for the follow up samples (P-value =
4.19e-10). Although there were discordent results the respective
sensitivity for the initial group was 100\% and for follow up was 100\%,
respectively.

There was 1 individual who still clearly had CRC on follow up as well as
5 individuals whose status on follow up was unknown. Although the 1
individual had a decrease in positive probability their follow up sample
was still higher than the cutoff threshold of 0.5 (positive probability
= 0.927). Interestingly, 2 individuals who were unknown on follow up
still were over the threshold cutoff of 0.5 even though, like the 1
individual with clear CRC on follow up, the probability of an adenoma or
carcinoma decreased {[}Table S3{]}.

The follow up positive probabilities were not affected by either
chemotherapy treatment (uncorrected P-value = 0.919) or radiation
therapy (uncorrected P-value = 1). There was also no difference in the
amount of change in the positive probability based on whether
individuals received chemotherapy (unccorected P-value = 0.578) or
radiation therapy (uncorrected P-value = 0.904).

\textbf{\emph{Specific OTUs in the Lesion Model are not Detected in
Follow Up Versus Initial Samples}} Overall, there were a total of 8 OTUs
that were common between the main lesion model and the model for
classifying initial and follow up samples specifically {[}Table S4{]}. A
total of 1 OTU was still significant after multiple comparison
correction and it's lowest taxonomic identification was to Blautia. In
general, Otu000012 (Blautia) was decreased from initial to follow up
{[}Figure 5{]}. The relative abundance was not drastically different
then the mean of the values observed in the control training set
{[}Figure 5{]}.

@@

We next created a model that incorporated both patient metadata, FIT,
and the bacterial microbiome to be able to predict lesions (adenoma or
carcinoma). We chose the middle training model, based on AUC, from 100
80/20 (train/test) splits. It's 10-fold cross validated AUC was similar
to it's test set AUC which was not the case for both the best and worse
training model {[}Figure 3{]}. Using this model we predicted the
probability of a lesion in the inital and follow up samples {[}Figure
4{]}. There was a signficant decrease in positive probability regardless
of whether the sample as a carcinoma or adenoma. The overall sensitivity
for lesion detection in the intial samples was 100 and for follow ups
was 100. Although there was a decrease in overall probability of an
adenoma or carcinoma only 24 were below the 0.5 threshold out of the
total 65 individuals who were diagnosed as not having a carcinoma on
follow up.

We then investigated which OTUs could potentially be more important in
our model {[}Figure 5 \& Table S4{]}. Only a single OTU was significant
after multiple comparison correction and the lowest taxonomic
identification of Otu000012 was to Blautia. Although there was a
difference in the relative abundance at initial and follow up these
values were not drastically different from the relative abundance values
observed in the control individuals of the training set {[}Figure 5{]}.
This research provides evidence that it is possible to create a highily
sensitive model for detection of adenoma or carcinoma with bacterial
microbiome data. It accomplishes this by using a unique sample set in
which before and after stool samples are available for assessment. By
using these types of samples we are able to show that this model is
reactive. That is to say that after surgery for removal of the adenoma
or carcinoma it decreases the positive probability to reflect a lower
likelihood of the individual having an adenoma or carcinoma.

@@

\textbf{\emph{Study Design and Patient Sampling}} The sampling and
design of the study was similar to that reported in Baxter, et al
(\textbf{insert citation}). In brief, study exclusion involved those who
had already undergone surgery, radiation, or chemotherapy, had
colorectal cancer before a baseline stool sample could be obtained, had
IBD, a known hereditary non-polyposis colorectal cacner, or Familial
adenomatous polyposis. Samples used to build the model used for
prediction were collected either prior to a colonoscopy or between 1 - 2
weeks after. The bacterial microbiome has been shown to nomralize within
this time period (\textbf{insert citation}). Kept apart from this
training set were a total of 67 individuals that not only had a sample
as described previoulsy but also a follow up sample between 188 - 546
days after surgery and treatment had been completed. This study was
approved by the University of Michigan Institutional Review Board. All
study participants provided informed consent and the study itself
conformed to the guidelines set out by the Helsinki Declaration.

@@

\textbf{\emph{Sequence Processing}} The mothur software package
(v1.37.5) was used to process the 16S rRNA gene sequences. This process
has been previously described (\textbf{insert citations}). The general
processing workflow using mothur is as follows: Paired-end reads were
first merged into contigs, quality filtered, aligned to the SILVA
database, screening for chimeras, classified with a naive Bayesian
classifier using the Ribosomal Database Project (RDP), and clustered
into Operational Taxonomic Units (OTUs) using a 97\% similarity cutoff
with an average neighbor clustering algorithm. The numer of sequences
for each sample was rarified to 10521 in an attempt to minimize uneven
sampling.

\textbf{\emph{Model Creation}} The Random Forest (\textbf{insert
citation}) algorithm was used to create the model used for prediction of
lesion (adenoma or carcinoma) for the 67 individuals with follow up
samples. The model included data on FIT, the bacterial microbiome, sex,
age, Body Mass Index (BMI), whether the indivdiual was caucasion or not,
history of cancer, and family history of cancer. Non-binary data was
checked for near zero variance and auto correlation. Data columns that
had near zero variance were removed. Columns that were correlated with
each other over a Spearman correlation coefficient of 0.75 had one of
the two columns removed. This pre-processing was performed with the R
package caret (v6.0.73). Optimiation of the mtry hyperparameter as well
as data on the best and worse performance of the model involved taking
the 490 samples and making 100 80/20 (train/test) splits in the data
where control and lesion were equally represented in the 80 and 20
split, respectively. This 80\% portion was then split again into an
80/20 split, and run twenty times through 10-fold cross validation to
optimzie the model's Receiver Operator Characteristic (ROC) and tuneing
both mtry and ntree. This resulting model was then tested on the 20\% of
the data that was originally held out from this overall process. This
was repeated 100 times with the best model after all the repeats chosen
to be used on the 67 samples with follow up. Once the ideal mtry was
found the entire 490 sample set was used to create the final Random
Forest model. The default cutoff of a probability of 0.5 was used as the
threshold to classify individuals as positive or negative of lesion.

@@

\textbf{\emph{Statistical Analysis}} The R software package (v3.3.0) was
used for all statisitical analysis. Comparisons between bacterial
community structure utilized PERMANOVA (\textbf{insert citation}) in the
vegan package (v2.4.1) while comparisons between ROC curves utilized the
method by DeLong et al. (\textbf{insert citation}) executed by the pROC
(v1.8) package. Comparisons between probabilities as well as overall
amount of OTU between initial and follow up samples utilized a paired
wilcoxson ranked sum test. Where multiple comparison testing was needed
a Benjamini-Hochberg (BH) correction was applied (\textbf{insert
citation}) and a corrected P-value of less than 0.05 was considered
significant. Unless otherwise stated the P-values reported are those of
the BH corrected ones.

@@

\textbf{Figure 1: Change in Thetayc and Fit between initial and follow
up in adenoma or carcinoma group.} A) No significant difference was
found between the adenoma and carcinoma group for thetayc (P-value =
0.697). B) A significant difference was found between the adenoma and
carcinoma group for Fit (P-value = 2.15e-05).

@@

\textbf{Figure 3: Graph of the Receiver Operating Characteristic Curve
on Test Set Performance of the Best, Middle, and Worse Training Models.}
For each of the 100 training cohort sets used had 392 individuals and
the testing cohort sets had 13 individuals. The AUC on the test sets for
the best, middle, and worse models from training were 0.682, 0.815, and
0.931, respectively. cvAUC is the 10-fold cross-validated AUC from
training.

@@

\textbf{Figure 5: Lesion Model OTU with a Significant Decrease in
Relative Abundance that is also Predictive of Initial and Follow Up.}
After multiple comparison correction 1 (Blautia) was the only one with a
P-value \textless{} 0.05. The dotted line represents the average
relative abundance in the control training group.
